---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geopipeline

<!-- badges: start -->
<!-- badges: end -->

The goal of geopipeline is to streamline the data access and processing of geospatial datasets,
particularly for modeling workflows. geopipeline makes it easy to produce 2 types of outputs:

1. **Data Frame** - Produced as an output when a user wants to extract the values associated with particular points in a region of interest. The data frame can return the pixel values at the exact point locations, or the pixel values of all pixels that fall within a specified radius of the point locations. The data frame format may be useful for training a model.  

2. **GeoTIFF File(s)** - Produced as an output when the user wants to extract the values associated with an entire region of interest. When a user defines an area of interest as a polygon, the output of the core geopipeline functions is a GeoTIFF file of the bounding box for the polygon.

There are currently 7 core functions in geopipeline that can return data frame or GeoTIFF outputs for 7 unique datasets:

1. `s2_process` - Sentinel-2 L2A
2. `s1_process` - Sentinel-1 Radiometrically Terrain Corrected 
3. `dem_process` - Copernicus DEM GLO-30
4. `esa_worldcover_process` - ESA WorldCover
5. `alos_fnf_process` - ALOS Forest/Non-Forest
6. `soil_process` - ISRIC Soil Grids
7. `gedi_process` - GEDI L3

I would like to specifically acknowledge and cite the `rsi` package from Permian-Global-Research (https://github.com/Permian-Global-Research/rsi/tree/main) which is the foundation for accessing, processing, and calculating indices for the first 5 datasets listed above. That package is truly incredible and has made working on this project much easier! Additional functionality I have built on top of `rsi` for the intended modeling use case of geopipeline includes: 

* Creating a data frame output for regions of interest based on point geometries
* Adding the ability to spatially mosaic neighboring image tiles when temporal compositing is not applied for both Sentinel-2 L2A data and Sentinel-1 RTC GRD data.

Still to come: 

* A quick function for stacking and labeling non-composited images covering the same spatial extent.
* A function for summarizing pixel values within a buffer based on a user-defined formula. 
* Core functions for accessing GEDI L2A and L2B gridded datasets.

## Installation

You can install the development version of geopipeline from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("laurenkwick/geopipeline")
```

## Example

Begin by loading example data. In this example, the region of interest is the George W Mead Wildlife Area in Wisconsin. The point data includes fires that recently occurred within the wildlife area.

```{r load_shp_files}
library(geopipeline)

sf_point <- sf::st_read("~/ProcessingModule/national_forest_data/mead_fires.shp")
sf_poly <- sf::st_read("~/ProcessingModule/national_forest_data/mead_boundary.shp", type = 3)
```
We will use s2_process to return a GeoTIFF for the region of interest and a data frame containing values for our specific points of interest. 

```{r run_composites}
# Polygon data
s2_composite_im <- s2_process(sf_poly, start_dt = "2024-07-01", end_dt = "2024-07-31",
                              layers_sel = c("red", "nir"), 
                              composite_method = "median", idx_names = "NDVI", 
                              file_path = tempfile(pattern="s2_composite_im", tmpdir=tempdir()))

# Point data
s2_composite_df <- s2_process(sf_point, start_dt = "2024-07-01", end_dt = "2024-07-31",
                              uniqueID = "FIRE_ID",
                              layers_sel = c("red", "nir"), 
                              composite_method = "median", idx_names = "NDVI",
                              file_path = tempfile(pattern="s2_composite_df", tmpdir=tempdir()))
```
Plot the image: 
```{r plot_composites}
terra::plot(terra::rast(s2_composite_im))
```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
