---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geopipeline

<!-- badges: start -->
<!-- badges: end -->

The goal of geopipeline is to streamline the data access and processing of geospatial datasets,
particularly for modeling workflows. geopipeline makes it easy to produce 2 types of outputs:

1. **Data Frame** - Produced as an output when a user wants to extract the values associated with particular points in a region of interest. The data frame can return the pixel values at the exact point locations, or the pixel values of all pixels that fall within a specified radius of the point locations. The data frame format may be useful for training a model.  

2. **GeoTIFF File(s)** - Produced as an output when the user wants to extract the values associated with an entire region of interest. When a user defines an area of interest as a polygon, the output of the core geopipeline functions is a GeoTIFF file of the bounding box for the polygon.

There are currently 7 core functions in geopipeline that can return data frame or GeoTIFF outputs for 7 unique datasets:

1. `s2_process` - Sentinel-2 L2A
2. `s1_process` - Sentinel-1 Radiometrically Terrain Corrected 
3. `dem_process` - Copernicus DEM GLO-30
4. `esa_worldcover_process` - ESA WorldCover
5. `alos_fnf_process` - ALOS Forest/Non-Forest
6. `soil_process` - ISRIC Soil Grids
7. `gedi_process` - GEDI L3

I would like to specifically acknowledge and cite the `rsi` package from [Permian-Global-Research](https://github.com/Permian-Global-Research/rsi/tree/main) which is the foundation for accessing, processing, and calculating indices for the first 5 datasets listed above. That package is truly incredible and has made working on this project much easier! Additional functionality I have built on top of `rsi` for the intended modeling use case of geopipeline includes: 

* Creating a data frame output for regions of interest based on point geometries
* Adding the ability to spatially mosaic neighboring image tiles when temporal compositing is not applied for both Sentinel-2 L2A data and Sentinel-1 RTC GRD data.

*Still to come:* 

* A quick function for stacking and labeling non-composited images covering the same spatial extent.
* A function for summarizing pixel values within a buffer based on a user-defined formula. 
* Core functions for accessing GEDI L2A and L2B gridded datasets.

## Installation

You can install the development version of geopipeline from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("laurenkwick/geopipeline")
```

## Example

Begin by loading example data. In this example, the region of interest is the George W Mead Wildlife Area in Wisconsin. The point data includes fires that recently occurred within the wildlife area. Data for this example was acquired from [GeoData@Wisconsin](https://geodata.wisc.edu/)

```{r load_shp_files}
library(geopipeline)

sf_point <- sf::st_read("~/ProcessingModule/national_forest_data/mead_fires.shp")
sf_poly <- sf::st_read("~/ProcessingModule/national_forest_data/mead_boundary.shp", type = 3)
```

### July - Temporal Compositing

We will use s2_process to return a GeoTIFF for the region of interest and a data frame containing values for our specific points of interest. 

```{r run_composites}
# Polygon data
s2_composite_im <- s2_process(sf_poly, start_dt = "2024-07-01", end_dt = "2024-07-31",
                              layers_sel = c("red", "nir"), 
                              composite_method = "median", idx_names = "NDVI", 
                              file_path = tempfile(pattern="s2_composite_im", tmpdir=tempdir()))

# Point data
s2_composite_df <- s2_process(sf_point, start_dt = "2024-07-01", end_dt = "2024-07-31",
                              uniqueID = "FIRE_ID", layers_sel = c("red", "nir"), 
                              composite_method = "median", idx_names = "NDVI",
                              file_path = tempfile(pattern="s2_composite_df", tmpdir=tempdir()))
```
Plot the image: 

```{r plot_composites, echo=FALSE}
# Plot
terra::plot(terra::rast(s2_composite_im))
```
View the data frame. In geopipeline, data frames are serialized as an fst file by default, because they can very quickly become very large. To read in the data frame, use `fst::read_fst()`: 

```{r df_composite}
# Data frame
s2_df_composite <- fst::read_fst(s2_composite_df)
head(s2_df_composite)
```

IDs in the data frame "ID" column correspond to the ID from our original shapefile attribute table, which we specified in the `s2_process` function using `uniqueID = "FIRED_ID"`. In the previous example, we have one pixel (or one row) per point in our original shapefile. In this next example, we will use the `radius` argument to grab all pixels that fall within a 100m distance to each point:

```{r composite_with_radius}
# Run s2_process
s2_composite_radius_df <- s2_process(sf_point, radius = 100, start_dt = "2024-07-01", end_dt = "2024-07-31",
                                     uniqueID = "FIRE_ID", layers_sel = c("red", "nir"), 
                                     composite_method = "median", idx_names = "NDVI",
                                     file_path = tempfile(pattern="s2_composite_radius_df", tmpdir=tempdir()))

# Read data frame
s2_df_composite_radius <- fst::read_fst(s2_composite_radius_df)
head(s2_df_composite_radius)
```
We can see that there are now several rows associated with FIRE_ID 170. 

### July - No Temporal Compositing

If we do not want apply a temporal composite over the time range, we can return all images that were acquired within the date range. If the region of interest spans over multiple Sentinel-2 L2A, geopipeline will still mosaic neighboring image tiles, as long as they were acquired on the same day. GeoTIFF files will be saved with the date of acquisition appended at the end. These files can be converted into raster data and stacked to create one large raster.

In the case of data frames, only one data frame is returned with each column of the data frame having an associated date of acquisition.

In this next example, we will run `s2_process` without applying a compositing function:

```{r no_composites}
# Polygon data
s2_no_composite_im <- s2_process(sf_poly, start_dt = "2024-07-01", end_dt = "2024-07-31",
                                 layers_sel = c("red", "nir"), 
                                 composite_method = NULL, idx_names = "NDVI", 
                                 file_path = tempfile(pattern="s2_no_composite_im", tmpdir=tempdir()))

# Point data
s2_no_composite_df <- s2_process(sf_point, start_dt = "2024-07-01", end_dt = "2024-07-31",
                                 uniqueID = "FIRE_ID", layers_sel = c("red", "nir"), 
                                 composite_method = NULL, idx_names = "NDVI",
                                 file_path = tempfile(pattern="s2_no_composite_df", tmpdir=tempdir()))
```

Choosing to not apply a compositing funciton may be useful if you want to perform complex time series analysis later on, one that takes seasonal variation into account. 

We will visualize the results now: 

```{r plot_no_composites, echo=FALSE}
# Plot
terra::plot(terra::rast(s2_no_composite_im[1]))
terra::plot(terra::rast(s2_no_composite_im[2]))
```

View the dataframe:

```{r df_no_composite}
# Data frame
s2_df_no_composite <- fst::read_fst(s2_no_composite_df)
head(s2_df_no_composite)
```


